\pkg{regsem}: Regularized Structural Equation Modeling

### 2017-03-07 draft1.Rmd
first full draft

This paper provides an introduction to the \pkg{regsem} package, outlining the mathematical details of the regularized SEM method and the usage of the package itself.

## Future Directions

## Limitations

With highly constrained structural equation models, achieving convergence can be particularly problematic in using \pkg{regsem}. For instance, with the latent change score model (cite jack), Bayesian regularization methods have less difficulty in reaching convergence across chains (cite chapter). With the advent of additional sparsity inducing priors, along with new forms of software such as STAN (cite stan), for some models it may be more appropriate to use these Bayesian regularization methods over their frequentist counterparts.

@@

>>>>

This paper provides an introduction to the \pkg{regsem} package, outlining the mathematical details of the regularized structural equation modeling [RegSEM; @jacobucci2016regularized] method and the usage of the package itself. RegSEM allows the use of regularization while keeping the structural equation model intact, adding penalization directly into the estimation of the model. The application of RegSEM was detailed using two example models: a latent growth curve model with 20 predictors of both the latent intercept and slope, along with a factor analysis with one latent factor. With the latent growth curve model, the small parameter to sample size ratio resulted in a larger number of false positives in using maximum likelihood estimation. In both the simulated example and the small simulation, the different types of regularization in \pkg{regsem} demonstrated better false positive and negative rates in comparison to maxiumum likelihood across sample sizes.

Broadly speaking, there is a growing amount of research into the integration between data mining methods and latent variable models. Specifically, beyond regularized SEM, this has taken form of item response theory and regularization [@sun2016latent], pairing both structural equation models with decision trees [@brandmaier2013], exploratory psychological network analaysis [e.g. @epskamp2016generalized], along with many others. The amount of pairing between methods that have generally been housed in separate camps will only increase into the future. This type of research will be facilitated by the general upsurge in the creation of open source software that gives users a general framework to test models. This was the motivation behind creating the \pkg{regsem} package in that users can estimate models ranging from simple factor analysis models, to latent longitudinal models with few to many time points, and finally to models with a large number of latent and observed variables. The use of regularization allows for the estimation of much larger structural equation models than before. Specifically, given that latent variables, as discussed in this article, are more commonly used in the social and behavioral sciences, sample sizes are generally not large. To estimate large models with sample sizes invites increasing amounts of bias, and as demonstrated with the simulated data in this paper, regularization can be used to reduce the complexity of the model, decreasing both the bias and variance.

With highly constrained structural equation models, achieving model convergence can be particularly problematic in using \pkg{regsem}. For instance, with the latent change score model [@mcardle2001latent], Bayesian regularization methods have less difficulty in reaching convergence across chains (Jacobucci and Grimm 2017). With the recent advent of additional sparsity inducing priors, along with new forms of software such as STAN [@carpenter2016stan], for some models it may be more appropriate to use these Bayesian regularization methods over their frequentist counterparts. Particularly with structural equation models, although some research exists [@feng2017bayesian], much more is warranted.

Future research with \pkg{regsem} should focus on a number of avenues. One is comparing the different forms of regularization, delineating which method may best best in which circumstances.  Additionally, as structural equation models become larger, with the advent of much larger datasets, speed will become a principal concern. Although 40 penalties in the models tested above can be run in a matter of seconds on a standard laptop, larger models can take much longer. To handle this, future implementation with \pkg{regsem} will test the inclusion of different types of optimization, specifically testing whether computation of the Hessian (or approximate Hessian) can reduce the number of optimization steps.

@@

This paper provided a brief overview on the use of the \pkg{regsem} package as an implementation of regularized structural equation model. Because structural equation modeling encompasses a wide array of latent variable models, the \pkg{regsem} package was created as a general package for including different forms of regularization into a host of latent variable models. Future updates to \pkg{regsem} will focus on decreasing the computational time of large latent variable models in order to provide an avenue of testing for researchers collecting larger and larger datasets. RegSEM is a method that operates at all ends of the data size spectrum: allowing for a reduction in complexity when the sample size is small, along with  dimension reduction in the presence of large data (both $N$ and $P$).

# References

### 2017-03-07 draft1.tex
first full draft

This paper provides an introduction to the \pkg{regsem} package,
outlining the mathematical details of the regularized SEM method and the
usage of the package itself.

\subsection{Future Directions}\label{future-directions}

\subsection{Limitations}\label{limitations}

With highly constrained structural equation models, achieving
convergence can be particularly problematic in using \pkg{regsem}. For
instance, with the latent change score model (cite jack), Bayesian
regularization methods have less difficulty in reaching convergence
across chains (cite chapter). With the advent of additional sparsity
inducing priors, along with new forms of software such as STAN (cite
stan), for some models it may be more appropriate to use these Bayesian
regularization methods over their frequentist counterparts.

\subsection*{Conclusion}\label{conclusion}
\addcontentsline{toc}{subsection}{Conclusion}

\hypertarget{refs}{}
\hypertarget{ref-browne2001}{}
Browne, Michael W. 2001. ``An Overview of Analytic Rotation in
Exploratory Factor Analysis.'' \emph{Multivariate Behavioral Research}
36 (1). Taylor \& Francis: 111--50.

@@

@@

@@

@@

>>>>

This paper provides an introduction to the \pkg{regsem} package,
outlining the mathematical details of the regularized structural
equation modeling (RegSEM; Jacobucci, Grimm, and McArdle 2016) method
and the usage of the package itself. RegSEM allows the use of
regularization while keeping the structural equation model intact,
adding penalization directly into the estimation of the model. The
application of RegSEM was detailed using two example models: a latent
growth curve model with 20 predictors of both the latent intercept and
slope, along with a factor analysis with one latent factor. With the
latent growth curve model, the small parameter to sample size ratio
resulted in a larger number of false positives in using maximum
likelihood estimation. In both the simulated example and the small
simulation, the different types of regularization in \pkg{regsem}
demonstrated better false positive and negative rates in comparison to
maxiumum likelihood across sample sizes.

Broadly speaking, there is a growing amount of research into the
integration between data mining methods and latent variable models.
Specifically, beyond regularized SEM, this has taken form of item
response theory and regularization (Sun et al. 2016), pairing both
structural equation models with decision trees (Brandmaier et al. 2013),
exploratory psychological network analaysis (e.g. Epskamp, Rhemtulla,
and Borsboom 2016), along with many others. The amount of pairing
between methods that have generally been housed in separate camps will
only increase into the future. This type of research will be facilitated
by the general upsurge in the creation of open source software that
gives users a general framework to test models. This was the motivation
behind creating the \pkg{regsem} package in that users can estimate
models ranging from simple factor analysis models, to latent
longitudinal models with few to many time points, and finally to models
with a large number of latent and observed variables. The use of
regularization allows for the estimation of much larger structural
equation models than before. Specifically, given that latent variables,
as discussed in this article, are more commonly used in the social and
behavioral sciences, sample sizes are generally not large. To estimate
large models with sample sizes invites increasing amounts of bias, and
as demonstrated with the simulated data in this paper, regularization
can be used to reduce the complexity of the model, decreasing both the
bias and variance.

With highly constrained structural equation models, achieving model
convergence can be particularly problematic in using \pkg{regsem}. For
instance, with the latent change score model (McArdle and Hamagami
2001), Bayesian regularization methods have less difficulty in reaching
convergence across chains (Jacobucci and Grimm 2017). With the recent
advent of additional sparsity inducing priors, along with new forms of
software such as STAN (Carpenter et al. 2016), for some models it may be
more appropriate to use these Bayesian regularization methods over their
frequentist counterparts. Particularly with structural equation models,
although some research exists (Feng, Wu, and Song 2017), much more is
warranted.

Future research with \pkg{regsem} should focus on a number of avenues.
One is comparing the different forms of regularization, delineating
which method may best best in which circumstances. Additionally, as
structural equation models become larger, with the advent of much larger
datasets, speed will become a principal concern. Although 40 penalties
in the models tested above can be run in a matter of seconds on a
standard laptop, larger models can take much longer. To handle this,
future implementation with \pkg{regsem} will test the inclusion of
different types of optimization, specifically testing whether
computation of the Hessian (or approximate Hessian) can reduce the
number of optimization steps.

\subsection{Conclusion}\label{conclusion}

This paper provided a brief overview on the use of the \pkg{regsem}
package as an implementation of regularized structural equation model.
Because structural equation modeling encompasses a wide array of latent
variable models, the \pkg{regsem} package was created as a general
package for including different forms of regularization into a host of
latent variable models. Future updates to \pkg{regsem} will focus on
decreasing the computational time of large latent variable models in
order to provide an avenue of testing for researchers collecting larger
and larger datasets. RegSEM is a method that operates at all ends of the
data size spectrum: allowing for a reduction in complexity when the
sample size is small, along with dimension reduction in the presence of
large data (both \(N\) and \(P\)).

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-brandmaier2013}{}
Brandmaier, Andreas M., Timo von Oertzen, John J. McArdle, and Ulman
Lindenberger. 2013. ``Structural Equation Model Trees.''
\emph{Psychological Methods} 18 (1): 71--86.

\hypertarget{ref-browne2001}{}
Browne, Michael W. 2001. ``An Overview of Analytic Rotation in
Exploratory Factor Analysis.'' \emph{Multivariate Behavioral Research}
36 (1). Taylor \& Francis: 111--50.

\hypertarget{ref-carpenter2016stan}{}
Carpenter, Bob, Andrew Gelman, Matt Hoffman, Daniel Lee, Ben Goodrich,
Michael Betancourt, Michael A Brubaker, Jiqiang Guo, Peter Li, and Allen
Riddell. 2016. ``Stan: A Probabilistic Programming Language.''
\emph{Journal of Statistical Software} 20.

@@

\hypertarget{ref-epskamp2016generalized}{}
Epskamp, Sacha, Mijke Rhemtulla, and Denny Borsboom. 2016. ``Generalized
Network Psychometrics: Combining Network and Latent Variable Models.''
\emph{ArXiv Preprint ArXiv:1605.09288}.

@@

\hypertarget{ref-feng2017bayesian}{}
Feng, Xiang-Nan, Hao-Tian Wu, and Xin-Yuan Song. 2017. ``Bayesian
Regularized Multivariate Generalized Latent Variable Models.''
\emph{Structural Equation Modeling: A Multidisciplinary Journal}. Taylor
\& Francis, 1--18.

@@

\hypertarget{ref-mcardle2001latent}{}
McArdle, John J, and Fumiaki Hamagami. 2001. ``Latent Difference Score
Structural Models for Linear Dynamic Analyses with Incomplete
Longitudinal Data.'' American Psychological Association.

@@

\hypertarget{ref-sun2016latent}{}
Sun, Jianan, Yunxiao Chen, Jingchen Liu, Zhiliang Ying, and Tao Xin.
2016. ``Latent Variable Selection for Multidimensional Item Response
Theory Models via L\_ \(\{\)1\(\}\) Regularization.''
\emph{Psychometrika} 81 (4). Springer: 921--39.
